---
title: "Gov 50 Final Project"
author: "Angela Chen"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---

## Proposal

Does the rating of a company on glassdoor have a positive correlation with the maximum salary they offer for their web development role? In this study, I plan to examine whether a higher rating of a company typically indicates that the company pays a higher maximum salary for their web development role. 

I hypothesize that a higher company rating does positively correlate with higher maximum salary given for web developers. Glassdoor is a website where current and former employees can anonymously review companies. Therefore, companies that are rated higher typically have better resources,  working conditions, treatment of employees, etc… Not only do employees tend to rate companies higher if they were paid better there, but also companies with better resources (and hence better ratings), typically pay a higher salary. 

My sample consists of 900 companies total, with 612 companies having complete data on their glassdoor rating and their maximum and minimum salary for a web developer. This data was found through web scraping from glassdoor. The explanatory interest is the rating of the company on glassdoor, which is found in 0.1 increments from 1 to 5. My outcome variable of interest is the maximum salary the company has listed for its web development role in USD. 

If the company rating is positively correlated with the maximum salary offered, then this would provide support for my hypothesis. If, on the other hand, I observe that a company’s rating has a negative impact or no impact on its maximum salary offered, then this would provide evidence against my hypothesis. Beyond examining company rating effects on maximum salary for a web developer, I hope to also examine its effects on the minimum salary offered. 


```{r}
library(tidyverse)
library(dplyr)
library(broom)
```

```{r}
salary_rating <- read_csv("data/web_dev.csv")
salary_rating
```
```{r}
salary_rating <- na.omit(salary_rating)

salary_rating |> 
  select(company_rating, max_salary)
```
I removed all the rows that contained missing values as most of those missing values were found in the column on company_rating. 

```{r}
ggplot(data = salary_rating,
       mapping = aes(x = max_salary)) +
  geom_histogram() + 
  labs(x = "Maximum Salary for Web Developer", 
       y = "Count")
```
```{r}
ggplot(data = salary_rating,
       mapping = aes(x = company_rating,
                     y = max_salary)) +
  geom_point() + geom_smooth() + 
  labs(x = "Company Rating", 
       y = "Maximum Salary")
```


From this scatterplot and the smooth line, we see that there is a slightly negative correlation between the rating of a company on Glassdoor and the maximum salary that they pay for their standard web development job. 
